Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach

### 2023-03-29 paper.tex
Fix a typo
Add an example with a ball

\subsection{A disk}
If we define $f(x, y) = x^2 + y^2$ then an inequation $f(x, y) \le R^2$ defines
a disk of radius $R$ with center at the origin. In this case $F_{ss}$ depends
only on distance $r$ from the origin:
\begin{equation*}
  F_{ss}(r) = \left\{
  \begin{array}{ll}
    \frac{4R^2}{r\sqrt{4R^2-r^2}} & \quad r < 2R \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}
A comparison of the algorithm described in \cref{sec:algo} applied to images
with different resolution with a theory is shown on \cref{fig:fss-disk}.
\begin{figure*}[!pt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-3x3.png}
    \label{fig:fss-disk-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-5x5.png}
    \label{fig:fss-disk-5x5}}
  \caption[]{Plot of surface-surface function for a disk with radius
    $R = 0.0334$. Resolution of the image varies from $100\times 100$ pixels to
    $3100\times 3100$.}
  \label{fig:fss-disk}
\end{figure*}

\subsection{Gaussian mixture}
\label{sec:gauss}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth, frame]{images/blob.png}
  \caption[]{An example of a ``blob'' generated by Gaussian mixture $f(x,y)$
    which consists of four Gaussian functions. White area is where
    $f(x,y) \le 3$.}
  \label{fig:blob}
\end{figure}
Take $N$ random values $a_i$, $b_i$ and $\sigma_i$. Let $f$ be
\begin{equation*}
  f(x,y) = \sum_{i=1}^N \frac{1}{\sigma_i} \exp(\frac{-(x-a_i)^2-(y-b_i)^2}{\sigma_i^2})
\end{equation*}
Inequality $f(x,y) \le T$ defines a ``blob'' like one which can be seen on
\cref{fig:blob}. We can use the algorithm developed in \cref{sec:fss-2d} to
compute precise values of $F_{ss}$ for this blob and compare them with values
obtained by using the fast algorithm in section \cref{sec:algo}. The result of
this comparison is on \cref{fig:fss-blob}.
\begin{figure*}[!pt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-3x3.png}
    \label{fig:fss-blob-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-5x5.png}
    \label{fig:fss-blob-5x5}}
  \caption[]{Plot of the surface-surface function for the blob on
    \cref{fig:blob}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.3$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-blob}
\end{figure*}

\subsection{Ellipses}
\label{sec:ellipses}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth, frame]{images/ellipses.png}
  \caption[]{An example of generated ellipses.}
  \label{fig:ellipses}
\end{figure}
Take $N$ random values $a_i$, $b_i$, $c_i$, $x^c_i$ and $y^c_i$ and define
\begin{equation*}
  f_i(x, y) = a_i(x-x^c_i)^2 + b(x-x^c_i)(y-y^c_i) + c_i(y-y^c_i)^2
\end{equation*}
with $a_ic_i - b_i^2 > 0$. Inequation
\begin{equation*}
  \min_i f_i(x, y) < A^2
\end{equation*}
defines $N$ ellipses like those which can be seen on \cref{fig:ellipses}. A
comparison between two algorithms is on \cref{fig:fss-ellipses}.
\begin{figure*}[!pt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-3x3.png}
    \label{fig:fss-ellipses-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-5x5.png}
    \label{fig:fss-ellipses-5x5}}
  \caption[]{Plot of the surface-surface function for ellipses on
    \cref{fig:ellipses}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.1$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-ellipses}
\end{figure*}

@@

Now, let's try a wider filter $H'$:
\begin{align}
  H'_{ij} &= \frac{1}{30} \left[
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & -24 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array}
    \right] \label{eq:filter-5x5-2d} \\
  H'_{ijk} &= \frac{1}{150} \left[
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array} ;
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & -124 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array} ;
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array}
    \right] \label{eq:filter-5x5-3d}
\end{align}
With this filter we get a different picture (\cref{fig:angles-5x5}). A
comparison of the fast algorithm with a wider filter kernel with a precise
calculation for a ``blob'' described in \cref{sec:gauss} is show on
\cref{fig:fss-blob-5x5}. A case with ellipses from \cref{sec:ellipses} is on
\cref{fig:fss-ellipses-5x5}.
\begin{figure*}[!pt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/angles-3x3.png}
    \label{fig:angles-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/angles-5x5.png}
    \label{fig:angles-5x5}}
  \caption[]{Estimation of $1/\sin\phi$ where $\phi$ is the angle between
    tangent lines to the interface by filters $H$ and $H'$.}
  \label{fig:angles}
\end{figure*}

@@

\bibliography{paper}

\end{document}

>>>>

\subsection{A disk and a ball}
Define $f(\bm{x}) = \sum_i x_i^2$. If $\bm{x} \in \mathbb{R}^2$ when an
inequation $f(\bm{x}) \le R^2$ defines a disk of radius $R$ with the center at
the origin. In this case $F_{ss}$ depends only on distance $r$ from the origin:
\begin{equation*}
  F_{ss}(r) = \left\{
  \begin{array}{ll}
    \frac{4R^2}{r\sqrt{4R^2-r^2}} & \quad r < 2R \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}
A comparison of the algorithm described in \cref{sec:algo} applied to images
with different resolution with a theory is shown on \cref{fig:fss-disk}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-3x3.png}
    \label{fig:fss-disk-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-5x5.png}
    \label{fig:fss-disk-5x5}}
  \caption[]{Plot of surface-surface function for a disk with radius
    $R = 0.334$. Resolution of the image varies from $100\times 100$ pixels to
    $3100\times 3100$.}
  \label{fig:fss-disk}
\end{figure*}

If $\bm{x} \in \mathbb{R}^3$ then an inequation $f(\bm{x}) \le R^2$ defines a
ball. In this case we provide a comparison (\cref{fig:fsss-ball}) of the
algorithm in \cref{sec:algo} with the precise algorithm in
\cref{sec:fsss-3d}. There is no known analytic solution for this case.
\begin{figure*}[!hpt]
  \centering
  \includegraphics[width=0.45\linewidth]{images/ball-sss.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{images/ball-sss-error.png}
  \caption[]{On the left: Plot of $F_{sss}(\bm{r_1}, \bm{r_2})$ function
    obtained by algorithm \cref{sec:algo} for a ball with radius
    $R = 0.3$. $\bm{r_1}$ is parallel to a vector $\bm{i} = (1, 0, 0)$ and
    $\bm{r_2}$ is parallel to a vector $\bm{j} = (0, 1, 0)$. Vectors $\bm{i}$
    and $\bm{j}$ constitute a two dimensional parameter space for $F_{sss}$. On
    the right: Relative error of computation. True values for $F_{sss}$ are
    obtained by algorithm in \cref{sec:fsss-3d}.}
  \label{fig:fsss-ball}
\end{figure*}

\subsection{Gaussian mixture}
\label{sec:gauss}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth, frame]{images/blob.png}
  \caption[]{An example of a ``blob'' generated by Gaussian mixture $f(x,y)$
    which consists of four Gaussian functions. White area is where
    $f(x,y) \le 3$.}
  \label{fig:blob}
\end{figure}
Take $N$ random values $a_i$, $b_i$ and $\sigma_i$. Let $f$ be
\begin{equation*}
  f(x,y) = \sum_{i=1}^N \frac{1}{\sigma_i} \exp(\frac{-(x-a_i)^2-(y-b_i)^2}{\sigma_i^2})
\end{equation*}
Inequality $f(x,y) \le T$ defines a ``blob'' like one which can be seen on
\cref{fig:blob}. We can use the algorithm developed in \cref{sec:fss-2d} to
compute precise values of $F_{ss}$ for this blob and compare them with values
obtained by using the fast algorithm in section \cref{sec:algo}. The result of
this comparison is on \cref{fig:fss-blob}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-3x3.png}
    \label{fig:fss-blob-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-5x5.png}
    \label{fig:fss-blob-5x5}}
  \caption[]{Plot of the surface-surface function for the blob on
    \cref{fig:blob}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.3$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-blob}
\end{figure*}

\subsection{Ellipses}
\label{sec:ellipses}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth, frame]{images/ellipses.png}
  \caption[]{An example of generated ellipses.}
  \label{fig:ellipses}
\end{figure}
Take $N$ random values $a_i$, $b_i$, $c_i$, $x^c_i$ and $y^c_i$ and define
\begin{equation*}
  f_i(x, y) = a_i(x-x^c_i)^2 + b(x-x^c_i)(y-y^c_i) + c_i(y-y^c_i)^2
\end{equation*}
with $a_ic_i - b_i^2 > 0$. Inequation
\begin{equation*}
  \min_i f_i(x, y) < A^2
\end{equation*}
defines $N$ ellipses like those which can be seen on \cref{fig:ellipses}. A
comparison between two algorithms is on \cref{fig:fss-ellipses}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-3x3.png}
    \label{fig:fss-ellipses-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-5x5.png}
    \label{fig:fss-ellipses-5x5}}
  \caption[]{Plot of the surface-surface function for ellipses on
    \cref{fig:ellipses}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.1$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-ellipses}
\end{figure*}

@@

Now, let's try a wider filter $H'$:
\begin{align}
  H'_{ij} &= \frac{1}{30} \left[
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & -24 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array}
    \right] \label{eq:filter-5x5-2d} \\
  H'_{ijk} &= \frac{1}{150} \left[
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array} ;
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & -124 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array} ;
    \begin{array}{ccccc}
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
      1 & 1 & 1 & 1 & 1 \\
    \end{array}
    \right] \label{eq:filter-5x5-3d}
\end{align}
With this filter we get a different picture (\cref{fig:angles-5x5}). A
comparison of the fast algorithm with a wider filter kernel with a precise
calculation for a ``blob'' described in \cref{sec:gauss} is show on
\cref{fig:fss-blob-5x5}. A case with ellipses from \cref{sec:ellipses} is on
\cref{fig:fss-ellipses-5x5}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/angles-3x3.png}
    \label{fig:angles-3x3}}
  \hfill
  \subfigure[$5\times 5$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/angles-5x5.png}
    \label{fig:angles-5x5}}
  \caption[]{Estimation of $1/\sin\phi$ where $\phi$ is the angle between
    tangent lines to the interface by filters $H$ and $H'$.}
  \label{fig:angles}
\end{figure*}

@@

\bibliography{paper}
\end{document}
